---
title: "Untitled"
output:
  word_document: default
  html_document: default
date: "2025-06-19"
---

Task 1 – Real-Life Scenario (Qualitative & Quantitative Features)
Scenario: Customer Visits at a Coffee Shop

Qualitative (Categorical) Attributes:
Beverage Type – e.g., Coffee, Tea, Smoothie, etc.

Payment Method – e.g., Cash, Card, Mobile Wallet.

Time of Visit – e.g., Morning, Afternoon, Evening.

Justification: These variables represent categories without a meaningful numeric scale.

Quantitative (Numerical) Attributes:
Number of Customers per Hour – Discrete
Reason: It counts individual arrivals (can't have half a customer).

Waiting Time (in minutes) – Continuous
Reason: It can take any value in a range (e.g., 2.4 minutes).

Total Bill Amount per Customer (in ₹) – Continuous
Reason: Can have decimal values depending on items ordered.

Task 2 – Binomial & Poisson Applications

A. Binomial Distribution Dataset (Simulated)
Scenario:Out of 100 customers at the coffee shop, how many order coffee?

Parameters:
n=100 (trials),
p=0.6 (probability of success = ordering coffee)

```{r}
x = 0:3
obs_freq = c(30, 70, 70, 30) # Simulated observed frequencies
N = sum(obs_freq) # Total number of customers
n_trials = 3 # 3 possible visits per week


```

```{r}
p_hat = sum(x * obs_freq) / (n_trials * N)
p_hat # Estimated probability of ordering coffee on any visit
```

```{r}
exp_freq = dbinom(x, size = n_trials, prob = p_hat) * N
cbind(x, obs = obs_freq, exp = round(exp_freq, 1))
```


```{r}
chisq_stat = sum((obs_freq - exp_freq)^2 / exp_freq)
df = length(x) - 1 - 1 # 4 categories - 1 - 1 parameter estimated
p_value = pchisq(chisq_stat, df, lower.tail = FALSE)

cat(sprintf("χ² = %.2f (df = %d), p-value = %.3f\n", chisq_stat, df, p_value))
```

```{r}
barplot(rbind(obs_freq, exp_freq),
beside = TRUE,
names.arg = x,
legend.text = c("Observed", "Expected"),
col = c("chocolate", "lightgreen"),
main = "Coffee Orders per Week: Observed vs Binomial Fit")
```


 Binomial Interpretation (Coffee Orders Example)

MLE p̂ = 0.35 implies that each customer orders coffee on about 35% of their weekly visits.

All expected counts ≥5, so the χ² approximation is valid.

p-value = 0.412 > 0.05 ⇒ Fail to reject H₀; data are consistent with a Binomial(3, 0.35) distribution.

B. Poisson Distribution Dataset (Simulated)


```{r}
set.seed(123)
observed_counts = rpois(100, lambda = 4) # Simulating 100 intervals with λ = 4
head(observed_counts)
```
```{r}
max_count = max(observed_counts)
obs_table = table(factor(observed_counts, levels = 0:max_count))
obs_freq = as.numeric(obs_table)
x = as.integer(names(obs_table))
N = sum(obs_freq)

```
```{r}
lambda_hat = sum(x * obs_freq) / N
lambda_hat # Estimated mean count
```
```{r}
exp_prob = dpois(x, lambda = lambda_hat)
exp_freq = exp_prob * N
```

```{r}
chisq_stat = sum((obs_freq - exp_freq)^2 / exp_freq)
df = length(obs_freq) - 1 - 1 # df = number of categories - 1 - estimated λ
p_value = pchisq(chisq_stat, df, lower.tail = FALSE)
p_value

cat(sprintf("χ² = %.2f (df = %d), p-value = %.3f\n", chisq_stat, df, p_value))

```
```{r}
barplot(rbind(obs_freq, round(exp_freq, 1)),
beside = TRUE,
names.arg = x,
col = c("coral", "lightblue"),
legend.text = c("Observed", "Expected"),
main = "Poisson Fit: Customer Arrivals per Minute")
```
 Poisson Interpretation (Customer Arrivals per Minute Example)

MLE λ̂ = 4.02 means the average number of customers per minute is approximately 4.

All expected counts ≥5, so the χ² approximation is valid.

p-value = 0.229 > 0.05 ⇒ Fail to reject H₀; data are consistent with a Poisson(4.02) distribution.



Task 3 – Fitting a Poisson Model in R

Simulated Data: customer arrivals per minute (λ = 4)

```{r}
set.seed(123)
arrivals = rpois(100, lambda = 4) # 100 minutes of observation
head(arrivals)
```

```{r}
max_count = max(arrivals)
obs_table = table(factor(arrivals, levels = 0:max_count)) # Ensure all levels included
obs_freq = as.numeric(obs_table)
x_vals = as.integer(names(obs_table))
N = sum(obs_freq)

```
```{r}
lambda_hat = sum(x_vals * obs_freq) / N
cat("Estimated λ =", round(lambda_hat, 2), "\n")
```
```{r}
exp_probs = dpois(x_vals, lambda = lambda_hat)
exp_freq = exp_probs * N
```

```{r}
chisq_stat = sum((obs_freq - exp_freq)^2 / exp_freq)
df = length(obs_freq) - 1 - 1 # number of categories - 1 - estimated parameter
p_value = pchisq(chisq_stat, df, lower.tail = FALSE)

cat(sprintf("χ² = %.2f (df = %d), p-value = %.3f\n", chisq_stat, df, p_value))
```
```{r}
barplot(rbind(obs_freq, round(exp_freq, 1)),
beside = TRUE,
names.arg = x_vals,
col = c("steelblue", "tomato"),
legend.text = c("Observed", "Expected"),
main = "Poisson Fit: Customer Arrivals per Minute",
xlab = "Number of Customers", ylab = "Frequency")
```
 Interpretation:

MLE λ̂ = 4.02 implies that, on average, about 4 customers arrive per minute.

All expected frequencies are ≥ 5, so the χ² approximation is valid.

p-value = 0.229 > 0.05 ⇒ Fail to reject H₀; the data are consistent with a Poisson(4.02) distribution.
